---
output:
  html_document:
    code_folding: null
    keep_md: false
    mathjax: null
    self_contained: false
    number_sections: no
    highlight: null
    toc: no
    includes:
      before_body: header.html
      after_body: footer.html
    theme: yeti
    md_extensions: -autolink_bare_uris
---


```{r buildXML_loadpkgs, echo=FALSE, message = FALSE, warnings=FALSE, tidy=TRUE}

library(leaflet)
library(xml2)
library(RODBC, quietly = TRUE, verbose = FALSE)
library(httr, quietly = TRUE, verbose = FALSE)

source("R/assign_fun.R", verbose = FALSE, echo = FALSE)
ezid_out <- suppressWarnings(assign_doi(ds_id, post = TRUE))

if (is.null(ezid_out[[2]])) {stop('Whoops')}

table <- read.csv('outputs.csv', header = TRUE, stringsAsFactors = FALSE)

table <- rbind(table, c(ds_id, ifelse(length(ezid_out[[2]]) > 1, "Err", ezid_out[[2]])))

write.csv(table, 'outputs.csv', row.names = FALSE)

# > 10.5072/FK20G3NT7Q

# Note, this is being run by a knit command in another document, so 
# ds_id is being passed in from the external script:

source('R/sql_calls.R')

con <- odbcDriverConnect('driver={SQL Server};server=db5.emswin.psu.edu\\MSSQLSERVER_2012;database=Neotoma;trusted_connection=true')
  
default <- sqlQuery(con, query = default_call(ds_id))
# Clean the affiliation:
default$affiliation <- gsub('\r\n', ', ', default$affiliation)

sharedds <- sqlQuery(con, query = sharedSite_call(ds_id))

dataset <- read_xml(paste0(end_point, '/', ds_id, '/', ds_id, '_output.xml'))

```

<title>`r default$SiteName[1]` - Neotoma Dataset `r ds_id`</title>

# Neotoma Dataset `r ds_id`

## `r default$SiteName[1]`

<div style="float: right;">
```{r leaflet_map, echo=FALSE, message = FALSE, warnings=FALSE, tidy=TRUE, fig.width=3, fig.height=3}

new_map <- leaflet() %>% addTiles()

# Now we have to pull out the lat/longs:
locs <- xml_text(xml2::xml_find_all(dataset, ".//d1:geoLocations", xml_ns(dataset)))
locs <-   strsplit(locs, split = ' ')[[1]]
locs <- as.numeric(locs[nchar(locs) > 0])

popup_content <- paste0("<b>",default$SiteName[1],"</b>",
                        "<br>View in <a href=http://apps.neotomadb.org//explorer/?datasetid=", 
                        ds_id, ">Explorer</a>")

if (locs[1] == locs[3] & locs[2] == locs[4]) {
  new_map <- addMarkers(new_map, 
                        popup = popup_content, 
                        lng = locs[2], lat = locs[1]) %>% 
    setView(lng = locs[2], lat = locs[1], zoom = 3)
} else {
  new_map <- addPolygons(new_map, 
                         lng = c(locs[c(2,4)]), 
                         lat = c(locs[c(1,3)]), popup = popup_content) %>%
    setView(lng = locs[2], lat = locs[1], 
            zoom = 3)
}

new_map
```
</div>

```{r, echo=FALSE, message = FALSE, warnings=FALSE, tidy=TRUE}
doi <- ezid_out[[2]]
  #xml_text(xml_find_all(dataset, ".//d1:identifier", xml_ns(dataset)))
```

**DOI**: [`r substr(doi, 5, nchar(doi))`](`r paste0("http://dx.doi.org/", substr(doi, 5, nchar(doi)))`)

**Links**: View on [Neotoma Explorer](http://apps.neotomadb.org/explorer/?datasetid=`r ds_id`) | [Raw JSON](http://api.neotomadb.org/v1/data/downloads/`r ds_id`)

**Constituent Database**: `r unlist(sqlQuery(con, query = constdb_call(ds_id)))`

**Dataset Type**: `r sharedds$DatasetType[which(sharedds$DatasetID == ds_id)]`

**Latitude**: `r mean(locs[1], locs[3])`

**Longitude**: `r mean(locs[2], locs[4])`

```{r, echo=FALSE, message = FALSE, warnings=FALSE, tidy=TRUE}
agerange <- sqlQuery(con, query = agerange_call(ds_id))
if (nrow(agerange) == 0) {
  range_out <- "No relevant chronology."
} else {
  range_out <- paste0(agerange[1], ' -- ', agerange[2], 'ybp, ', agerange[3])
}
```

**Age Range**: 4700 - 7400 ybp

**Description**: `r unlist(sqlQuery(con, query = sitedesc_call(ds_id)))`

<br>

## Affiliated Researchers

```{r getContacts, results='as-is', echo=FALSE, message = FALSE, warnings=FALSE, tidy=TRUE}

contacts <- sqlQuery(con, query = contributor_call(ds_id))
contacts <- apply(contacts, 2, function(x){ iconv(x, "UTF-8", 'latin1')})

contacts[,"affiliation"] <- gsub('\r\n', ', ', contacts[,"affiliation"])
colnames(contacts) <- c("Name", "Affiliation", "Contribution")

contacts <- contacts[order(contacts[,1]), ]
contacts <- contacts[!duplicated(contacts),]

knitr::kable(contacts, row.names = FALSE)

```


## Publications

```{r, echo=FALSE, message = FALSE, warnings=FALSE, tidy=TRUE, results='as-is'}
publications <- sqlQuery(con, query = pub_call(ds_id), as.is = TRUE)
#publications[,4] <- iconv(publications[,4], "UTF-8", "latin1")

if (nrow(publications) > 0) {
  knitr::kable(publications[,c(1,4)])
} else {
  cat("No publications in this record.")
}

```

## Data Citation

```{r, echo=FALSE, message = FALSE, warnings=FALSE, tidy=TRUE}

plain_text <- paste0(contacts[which(contacts[,"Contribution"] == "ProjectLeader"),1], " ",
       xml_text(xml2::xml_find_all(dataset, ".//d1:publicationYear", xml_ns(dataset))),
       ". ",
       default$SiteName[1], ". In, *",
       unlist(sqlQuery(con, query = constdb_call(ds_id))), "*. ",
       "Neotoma Paleoecological Database. ", ezid_out[[2]],"<br>")
       
bibtex <- paste0("@inbook{<br>&nbsp;&nbsp;author={",
                 contacts[which(contacts[,"Contribution"] == "ProjectLeader"),1],
                 "},<br>&nbsp;&nbsp;year={",
                 xml_text(xml2::xml_find_all(dataset, ".//d1:publicationYear", xml_ns(dataset))),
                 "},<br>&nbsp;&nbsp;title={",
                 default$SiteName[1], "},<br>&nbsp;&nbsp;booktitle={",
                 unlist(sqlQuery(con, query = constdb_call(ds_id))), 
                 "},<br>&nbsp;&nbsp;publisher={",
                 "Neotoma Paleoecological Database},<br>&nbsp;&nbsp;doi={",
                 ezid_out[[2]],"}<br>}<br>")

```

### APA

`r plain_text`

### BibTeX

`r bibtex`

## Other Associated Datasets

### Other Datasets at Same Site:
```{r, results = 'as-is', echo=FALSE, message = FALSE, warnings=FALSE, tidy=TRUE}

sharedds <- sharedds[-which(sharedds$DatasetID == ds_id),c("Dataset", "DatasetID")]

if (nrow(sharedds) > 0) {
  sharedds$JSON <- paste0("[link](http://api.neotomadb.org/v1/data/downloads/",sharedds[,2],")")
  sharedds$Explorer <- paste0("[link](http://apps.neotomadb.org/explorer/?datasetid=",sharedds[,2],")")
  knitr::kable(sharedds[,-2], row.names = FALSE)
} else {
  cat("No other datasets at this site.")
}

odbcCloseAll()
```

### Most recent Neotoma Database snapshot: 

[http://www.neotomadb.org/snapshots](http://www.neotomadb.org/snapshots)

## Neotoma Data Use Policy 

We ask all data users to consider the Neotoma Data Use policy -- [link](http://www.neotomadb.org/data/category/use) -- as well as the general guidelines of open and ethical data sharing when using this data.